E Coli Project Paper Incredibly Rough Draft

### 2013-10-09 paper/paper.tex
added writing and plot Tested everything on node22.

%% \begin{figure}
%%   \includegraphics[width=\columnwidth]{../data/shape-p/plots/frequency-plot-all-NflD-p-40-20-0-0-150.pdf}
%%   \caption{Protein density fluctuation with time against the polar
%%     wall in a 4$\mu m$ by 2$\mu m$ pill shaped cell.  The wall-bound
%%     proteins are shown at a point that is on the wall and the
%%     cytoplasm densities are shown at a point immediately adjecent.
%%     This plot is normalized so that every protein density peaks at 1}
%%   \label{frequency-plot-40-20-0-0-150}
%% \end{figure}

@@

%% \begin{figure*}
%%   \includegraphics[width=\textwidth]{../data/shape-p/plots/image-plot_E--p-200-50-0-0-1500}
%%   \caption{Frames from an animation showing a series of contour plots
%%     of the concentration of proteins in different stages over time.
%%     The density of proteins are integrated along the axis
%%     normal to the page. The frames occur 2.5 seconds apart.}
%%   \label{total-oscillation-plot}
%% \end{figure*}

This simple model is a good starting point for observing in
detail the dynamic interaction between the different stages of protein
that lead to their qualitatively distinct behavoir. Figure (Figure
image) shows the process in a series of frames taken from simulation
animations.  The causal chain is cyclical so we can begin our
observations at any stage in the cycle.  We look first at MinD:ATP
floating freely in the cytoplasm.  The protein will not react with any
other floating protein, and it will not spontenously shift back into
the MinD:ADP stage of the cycle.  It will diffuse until it attaches to
a cell membrane.  When compared to a protein that is in the middle
cylindrical section of the cell, a protein that is in an end cap
section is more likely to bump into a wall in a given number of
diffusive steps, since there is a greater cross section of possible
movement that is covered by a nearby wall.  All else being equal,
proteins will collect with a higher rate on the caps.  This effect is
made more pronounced by the MinD:ATP tendency to attract and cluster.
When adjacent to the membrane, the rate of binding to the membrane is
porportional to the amount of MinD already bound there.  Initially,
then, there is a strong tendency of MinD to cluster in the polar
regions of the cell.

The next observed mechanism is the formation of MinE 'rings' on the
membranes of the polar caps.  As the freely
diffusing MinE enters the cap portion it begins to encounter
concentrated membrane bound MinD.  As it bumps into the walls here it
binds to this MinD and remians for a time on the walls in compound
form.  When it is released, it can diffuse naturally outward or inward
towards the cell.  If it diffuses inward, there is less cross
sectional wall for it to bump into and also less concentrated MinD:ATP
to react with.  If it diffuses outward, the curvature of the cell wall
cuases there to be a greater chance that it will bump into the cell
membrane (for the same reason outlined for the MinD:ATP above) and if
there is MinD:ATP bound there, there is a chance that it will bind to
that protein and react again.  If it does diffuse in this directin, it
will likely react with MinD:ATP before getting very far into the
encap.  Thus after dettaching from the membrane at the edge of the
MinD:ATP encap cluster, the MinE is most likely to travel a small
distance further into the polar region and then attach and react again
with MinD:ATP protein there.  This process results in the observed
qualitative 'encroaching MinE ring' behavoir that released MinD:ADP
back into the cytoplasm.  This MinD:ADP will diffuse for a time before
spontanosly uindergowing hydrolysis \fixme{is this right?} and
reaching the MinD:ATP state that can bind to the membrane again.
There is a strong chance that it has by this time diffused out of the
polar region and is on its way to the opposite polar region of the
cell.  If it hasn't then it will reattach to the MinD:ATP on the cell
wall deeper into the endcap and will eventually be converged upon by
the MinE ring, only to be released again in MinD:ADP form.

@@

\begin{figure}
  %%\includegraphics[width=\columnwidth]{../data/shape-p/plots/box-plot_D--p-400-50-0-0-1500}
  \includegraphics[width=\columnwidth]{../data/shape-p/plots/box-plot_D--p-200-50-0-0-1500}
  \caption{Total protein fluctuation in the right, middle, and left
    parts of a 5$\mu m$ by 1$\mu m$ pill shaped cell (above) and a
    3$\mu m$ by 1$\mu m$ pill shaped cell (below).  The vertical axis shows
    stacked the total number of proteins that are of four different
    compound stages and in the right section of the cell (bottom four
    colors), in the middle of the cell (middle four colors) and in the
    left section of the cell (top four colors). The different forms of
    protein changes while the total number of MinD protein in the cell
    remains constant, shown by the fact that the very top line is
    horizontal.}
  \label{total-oscillation-plot}
\end{figure}

Figure \ref{total-oscillation-plot} is a plot of the total fraction of
MinD found in each stage of the cycle in each of three sections of the
cell (we have split the pill into two polar sections and one center
section) over time.  One complete oscillation period is shown, in
which one of the poles begins and ends at a maximum density
concentration and the other a minimum denisty concentration.  Looking
at the evolution of the latter, we see an initial high concentration
of cytoplasmic MinD:ATP, as the this protein diffuses into the polar
region, followed by a dip in this and large spike in membrane bound
MinD:ATP, as the protein clusters onto the polar cell wall, followed
by a spike in the memebrane bound MinE:MinD:ATP stage of protein,
together with a slight rise in cytoplasmic MinD:ADP as the MinD
dettaches from the wall.

@@

Figure \ref{frequency-plot-40-20-0-0-150} also shows that the middle
density oscillates at a frequency that is twice that of the polar
regions, which is to be expected considering the symmetry of the
center compared to the that of the polar regions.  Also, the
difference between density maxima and minima each oscillation is
smaller for the middle points than for the polar points.  This is
evident from simply watching the simulation movies - the fluctulations
are more extreme in the polar regions.

>>>>

@@

\begin{table}
  \begin{tabular}{|r|c|c|c|l|}
    \hline
    Length($\mu$) & 2.50 & 3.00 & 4.00 & 5.00\\
    \hline
    Period(sec) & sim & 33 & 38 & 48 \\ \cline{2-2}
    \hline
  \end{tabular}
  \caption{Period According to Pill Length}
  \label{table-p}
\end{table}

This simple model is a good starting point for observing in detail the
dynamic interaction between the different stages of protein that lead
to their qualitatively distinct behavoir. Figure \ref{image-p} shows
the process in a series of frames taken from simulation animations.
Each frame is 2.5 seconds ahead of the last.  The causal chain is
cyclical so we can begin our observations at any stage in the cycle.
We consider first MinD:ATP floating freely in the cytoplasm.  The
protein will not react with any other floating protein, and it will
not spontenously shift back into the MinD:ADP stage of the cycle.  It
will diffuse until it attaches to a cell membrane.  When compared to a
protein that is in the middle cylindrical section of the cell, a
protein that is in an end cap section is more likely to bump into a
wall in a given number of diffusive steps, since there is a greater
cross section of possible movement that is covered by a nearby wall.
All else being equal, proteins will collect with a higher rate on the
caps.  This effect is made more pronounced by the MinD:ATP tendency to
attract and cluster.  When adjacent to the membrane, the rate of
binding to the membrane is porportional to the amount of MinD already
bound there.  Initially, then, there is a strong tendency of MinD to
cluster in the polar regions of the cell.

The next observed mechanism is the formation of MinE 'rings' on the
membranes of the polar caps.  As the freely
diffusing MinE enters the cap portion it begins to encounter
concentrated membrane bound MinD.  As it bumps into the walls here it
binds to this MinD and remians for a time on the walls in compound
form.  When it is released, it can diffuse naturally outward or inward
towards the cell.  If it diffuses inward, there is less cross
sectional wall for it to bump into and also less concentrated MinD:ATP
to react with.  If it diffuses outward, the curvature of the cell wall
cuases there to be a greater chance that it will bump into the cell
membrane (for the same reason outlined for the MinD:ATP above) and if
there is MinD:ATP bound there, there is a chance that it will bind to
that protein and react again.  If it does diffuse in this directin, it
will likely react with MinD:ATP before getting very far into the
encap.  Thus after dettaching from the membrane at the edge of the
MinD:ATP encap cluster, the MinE is most likely to travel a small
distance further into the polar region and then attach and react again
with MinD:ATP protein there.  This process results in the observed
qualitative 'encroaching MinE ring' behavoir that releases MinD:ADP
back into the cytoplasm.  This MinD:ADP will diffuse for a time before
spontanosly uindergowing hydrolysis \fixme{is this right?} and
reaching the MinD:ATP state that can bind to the membrane again.
There is a strong chance that it has by this time diffused out of the
polar region and is on its way to the opposite polar region of the
cell.  If it hasn't then it will reattach to the MinD:ATP on the cell
wall deeper into the endcap and will eventually be converged upon by
the MinE ring, only to be released again in MinD:ADP form.

@@

\begin{figure*}
  \includegraphics[width=\textwidth]{../data/shape-p/plots/image-plot--p-400-50-0-0-1500.pdf}
  \caption{Frames from an animation showing a series of contour plots
    of the concentration of proteins in different stages over time.
    The density of proteins are integrated along the axis
    normal to the page. The frames occur 2.5 seconds apart.}
  \label{image-p}
\end{figure*}

Consider Figure \ref{image-p}.  At 0 seconds we can see cytoplasmic
nATP roughly evenly distributed throughout the cell.  There is nADP
concentrated in the left pole and it will transform into nATP.  At 2.5
seconds we see membrane bound MinD:ATP form on the right side of the
cell.  It avoids the left side because there is a concentration of
cytoplasmic MinE on the left side and compared to the right side,
which is empty of MinE, MinD:ATP attaching to the wall on the left has
no time to collect and cluster before reacting.  One can see how as
the MinE spreads out rightward, the the clustered membrane bound
MinD:ATP stays just at the edge of it, always to the right.  Where the
two meet there is at every time step a spike in membrane bound
MinD-MinE:ATP, resulting from the MinE reaction with the MinD on the
wall.  One can see from roughly 5 seconds to 17.5 seconds the
cytoplasmic MinE chases in a way the MinD towards the right, reacting
with it as it clusters on the wall.  Upon reaching the right side, the
MinD is in a way trapped, and the MinE now concentrated in that pole
will react until the MinD has diffused back towards the other side and
there is no more membrane bound MinD:ATP to react with.  The cycle
then begins again.

\begin{figure}
  %%\includegraphics[width=\columnwidth]{../data/shape-p/plots/box-plot_D--p-400-50-0-0-1500}
  \includegraphics[width=\columnwidth]{../data/shape-p/plots/box-plot_D--p-200-50-0-0-1500}
  \caption{Total protein fluctuation in the right, middle, and left
    parts of a 5$\mu m$ by 1$\mu m$ pill shaped cell (above) and a
    3$\mu m$ by 1$\mu m$ pill shaped cell (below).  The vertical axis shows
    stacked the total number of proteins that are of four different
    compound stages and in the right section of the cell (bottom four
    colors), in the middle of the cell (middle four colors) and in the
    left section of the cell (top four colors). The different forms of
    protein changes while the total number of MinD protein in the cell
    remains constant, shown by the fact that the very top line is
    horizontal.}
  \label{box-p}
\end{figure}

Figure \ref{box-p} is a plot of the total fraction of
MinD found in each stage of the cycle in each of three sections of the
cell (we have split the pill into two polar sections and one center
section) over time.  One complete oscillation period is shown, in
which one of the poles begins and ends at a maximum density
concentration and the other a minimum density concentration.  Looking
at the evolution of the latter, we see an initial high concentration
of cytoplasmic MinD:ATP, as the this protein diffuses into the polar
region, followed by a dip in this and large spike in membrane bound
MinD:ATP, as the protein clusters onto the polar cell wall, followed
by a spike in the memebrane bound MinE:MinD:ATP stage of protein,
together with a slight rise in cytoplasmic MinD:ADP as the MinD
dettaches from the wall.

@@

Figure also shows that the middle
density oscillates at a frequency that is twice that of the polar
regions, which is to be expected considering the symmetry of the
center compared to the that of the polar regions.  Also, the
difference between density maxima and minima each oscillation is
smaller for the middle points than for the polar points.  This is
evident from simply watching the simulation movies - the fluctulations
are more extreme in the polar regions.
